---
title: "PS08-Midterm Review"
author: "Sejal Nimkar"
date: "2024-12-12"
output:
  pdf_document:
    latex_engine: xelatex
---

## Toss a fair coin until we get 1 head and 2 tails. Random variables: \(X\) counts the number of tosses, U assigns 1 if the number of tosses is greater than 3, and 0 otherwise, \(W = U1 + U2 + 路 路 路 + U16\) where each Ui has the same distribution as U for i = 1, . . . , 16 and \(U_i\) is independent of \(U_j\) for \(i 谈 = j\).

  **a. What is the sample space for this experiment? Explain and share at least 3 possible outcomes.**
  
  Ans:The sample space for this experiment includes all possible sequences of coin tosses that result in exactly 1 head and 2 tails. For example:
  
- HTT
- THT
- TTH

Each sequence represents a distinct outcome. The sample space is the set of all such sequences.

  
\vspace{2cm}

  **b. What is the range of \(X\) ? Is it countable? Clearly explain.**
    
  Ans: The random variable \( X \), which counts the number of tosses required to achieve 1 head and 2 tails, has a range of values starting from 3 (the minimum number of tosses required) to infinity (if the process continues indefinitely). This range is countable since it consists of discrete integers.

  
\vspace{2cm}
  
  **c. Find \(P(U = 1)\) (write this as a fraction), the median, and the IQR of \(U\) . Show your work.**
  
  Ans: To find \( P(U = 1) \), note that \( U = 1 \) occurs when the number of tosses \( X \) is greater than 3. The probability of \( X \leq 3 \) can be computed by enumerating the sequences that result in exactly 1 head and 2 tails within 3 tosses:
\(
P(X \leq 3) = P(\text{HTT}) + P(\text{THT}) + P(\text{TTH}) = 3 \cdot \left(\frac{1}{2}\right)^3 = \frac{3}{8}.
\)
Thus,
\(
P(U = 1) = 1 - P(X \leq 3) = 1 - \frac{3}{8} = \frac{5}{8}.
\)

The median of \( U \) is \( 0.5 \), and the interquartile range (IQR) is \( 1 - 0 = 1 \).

  
\vspace{2cm}

  **d. Find \(EW , Var(W) \), and present the R code to get \(P(3 < W < 7)\) .**
  
  Ans: To compute \( \mathbb{E}[W] \) and \( \text{Var}(W) \):
\(
\mathbb{E}[U] = P(U = 1) = \frac{5}{8}, \quad \mathbb{E}[W] = 16 \cdot \mathbb{E}[U] = 16 \cdot \frac{5}{8} = 10.
\)
\(
\text{Var}(U) = P(U = 1)(1 - P(U = 1)) = \frac{5}{8} \cdot \frac{3}{8} = \frac{15}{64}, \quad \text{Var}(W) = 16 \cdot \text{Var}(U) = 16 \cdot \frac{15}{64} = \frac{240}{64} = 3.75.
\)

The R code to calculate \( P(3 < W < 7) \) is:

```{r}
p_w <- pbinom(6, size = 16, prob = 5/8) - pbinom(3, size = 16, prob = 5/8)
p_w

```
  
\vspace{2cm}

  **e. Let \(V = \sum_{i=1}^{16} (-1)^i \cdot U_i\). Assume that you get 105 realizations of W and 105 realizations of V and the results are stored in R vectors vec.w and vec.v, respectively. Using the code below, do you expect res1 to be bigger than res2? Clearly explain.**
  
res1 <- sum((vec.w - mean(vec.w))^2)
res2 <- sum((vec.v - mean(vec.v))^2)
  
  Ans: We can check the answer using simulations:
  
```{r}
set.seed(123)
n <- 10^5
vec.w <- rbinom(n, size = 16, prob = 5/8)
vec.v <- cumsum((-1)^(1:16) * vec.w)
res1 <- sum((vec.w - mean(vec.w))^2)
res2 <- sum((vec.v - mean(vec.v))^2)
res1 > res2


```
  
\vspace{2cm}
\newpage

## Imagine throwing darts at a dart board \(B = \{(x, y) : 0 \leq x \leq y \leq 2\}\). Suppose that the darts are thrown in such a way that they are certain to hit a point in \(B\), and that each point in \(B\) is equally likely to be hit. Define the random variable \(X\) to be the value of the \(x\)-coordinate of the point that is hit, and define the random variable \(Y\) to be the value of the \(y\)-coordinate of the point that is hit.

  **a. Draw and write down the PDF of \(X\) **
  
  Ans:The probability density function (PDF) of \(X\) can be found from the uniform distribution over the region \(B\). The length of the line \(y = x\) to \(y = 2\) varies with \(x\), so the height of the distribution will be proportional to this length.

The PDF of \(X\) is given by:

\(
f_X(x) = \begin{cases} 
2 - x & \text{if } 0 \leq x \leq 2, \\
0 & \text{otherwise}.
\end{cases}
\)
  
\vspace{2cm}

  **b. Find \(E[X]\), \(P(X < \mu)\), \(q_2\), and \(q_3\)**
  
  Ans:
  
  1. \(E[X]\):
\(
E[X] = \int_{0}^{2} x \cdot f_X(x) \, dx = \int_{0}^{2} x(2 - x) \, dx
\)
\(
= \int_{0}^{2} (2x - x^2) \, dx = \left[ x^2 - \frac{x^3}{3} \right]_{0}^{2} = \left(4 - \frac{8}{3}\right) = \frac{4}{3}.
\)

2. \(P(X < \mu)\): Since \(\mu = E[X] = \frac{4}{3}\),
\(
P(X < \mu) = \int_{0}^{\frac{4}{3}} f_X(x) \, dx = \int_{0}^{\frac{4}{3}} (2 - x) \, dx
\)
\(
= \left[ 2x - \frac{x^2}{2} \right]_{0}^{\frac{4}{3}} = 2 \cdot \frac{4}{3} - \frac{\left(\frac{4}{3}\right)^2}{2} = \frac{8}{3} - \frac{8}{9} = \frac{16}{9}.
\)

3. \(q_2\) (median): Solve for \(q_2\) such that:
\(
\int_{0}^{q_2} f_X(x) \, dx = 0.5.
\)
\(
\int_{0}^{q_2} (2 - x) \, dx = 0.5 \implies \left[ 2x - \frac{x^2}{2} \right]_{0}^{q_2} = 0.5.
\)
\(
2q_2 - \frac{q_2^2}{2} = 0.5 \implies q_2^2 - 4q_2 + 1 = 0.
\)
Solving this quadratic equation:
\(
q_2 = 2 - \sqrt{3}.
\)

4. \(q_3\) (75th percentile): Solve for \(q_3\) such that:
\(
\int_{0}^{q_3} f_X(x) \, dx = 0.75.
\)
\(
2q_3 - \frac{q_3^2}{2} = 0.75 \implies q_3^2 - 4q_3 + 1.5 = 0.
\)
Solving this quadratic equation:
\(
q_3 = 2 - \sqrt{2.5}.
\)

  
\vspace{2cm}

  **c.  Provide an example showing that \(X\) and \(Y\) are not independent. **
  
  Ans: If \(X\) and \(Y\) were independent, then \(P(Y \leq y | X = x) = P(Y \leq y)\). However, given that \(x \leq y \leq 2\), the conditional probability depends on \(x\). For example:
\(
P(Y \leq 1 | X = 0.5) \neq P(Y \leq 1).
\)

Specifically:
\(
P(Y \leq 1 | X = 0.5) = \frac{\text{Length of } y = 0.5 \text{ to } y = 1}{\text{Length of } y = 0.5 \text{ to } y = 2} = \frac{1 - 0.5}{2 - 0.5} = \frac{1}{3}.
\)
  
\vspace{3cm}

  **d. Let the dart board be \(B = \{(x, y) : 0 \leq x \leq y \leq c\}\). Write down the PDF of \(X\) in terms of \(c\). **
  
  Ans: The PDF of \(X\) for a general \(c > 0\) is:
\(
f_X(x) = \begin{cases} 
\frac{c - x}{\frac{c^2}{2}} & \text{if } 0 \leq x \leq c, \\
0 & \text{otherwise}.
\end{cases}
\)
The normalization factor is the area of the dartboard, which is \(\frac{c^2}{2}\).
  
\vspace{2cm}
\newpage

## Buses pass by my stop at the top of the hour, 10 minutes past, and 30 minutes past the hour (e.g., at 8:00, 8:10, 8:30, 9:00, 9:10, 9:30, etc.). I arrive at the stop at a completely random time during the day. If the random variable \(Y\) represents the time I wait for the bus, we solve the following:

 **a. Draw and write down the PDF for \(Y\).**
 
 Ans: The waiting time \(Y\) is distributed over the intervals defined by the bus schedule. Since buses arrive at 0, 10, and 30 minutes past the hour, the waiting time intervals are:
- \(0 \leq Y \leq 10\) with slope increasing from \(0\) to \(1/10\),
- \(10 \leq Y \leq 30\) with slope constant at \(1/10\),
- \(30 \leq Y \leq 60\) with slope decreasing from \(1/10\) to \(0\).

The PDF of \(Y\) is:
\(
f_Y(y) = 
\begin{cases} 
\frac{1}{10}y & \text{if } 0 \leq y \leq 10, \\
\frac{1}{10} & \text{if } 10 < y \leq 30, \\
\frac{6 - y}{30} & \text{if } 30 < y \leq 60, \\
0 & \text{otherwise}.
\end{cases}
\)
 
 \vspace{2cm}
 
 **b. Find the following. **
 
 Ans:
 
1. The expected waiting time:
 
The expected value of \(Y\) is:
\(
E[Y] = \int_{0}^{10} y \cdot \frac{1}{10}y \, dy + \int_{10}^{30} y \cdot \frac{1}{10} \, dy + \int_{30}^{60} y \cdot \frac{6 - y}{30} \, dy.
\)

For \(0 \leq Y \leq 10\):
\(
\int_{0}^{10} y \cdot \frac{1}{10}y \, dy = \frac{1}{10} \int_{0}^{10} y^2 \, dy = \frac{1}{10} \left[ \frac{y^3}{3} \right]_{0}^{10} = \frac{1}{10} \cdot \frac{1000}{3} = \frac{100}{3}.
\)

For \(10 \leq Y \leq 30\):
\(
\int_{10}^{30} y \cdot \frac{1}{10} \, dy = \frac{1}{10} \int_{10}^{30} y \, dy = \frac{1}{10} \left[ \frac{y^2}{2} \right]_{10}^{30} = \frac{1}{10} \left( \frac{900}{2} - \frac{100}{2} \right) = \frac{1}{10} \cdot 400 = 40.
\)

For \(30 \leq Y \leq 60\):
\(
\int_{30}^{60} y \cdot \frac{6 - y}{30} \, dy = \frac{1}{30} \int_{30}^{60} (6y - y^2) \, dy = \frac{1}{30} \left[ 3y^2 - \frac{y^3}{3} \right]_{30}^{60}.
\)
Substituting limits:
\(
\frac{1}{30} \left( 3(60^2) - \frac{60^3}{3} - 3(30^2) + \frac{30^3}{3} \right) = \frac{1}{30} \cdot 540 = 18.
\)

Adding these, \(E[Y] = \frac{100}{3} + 40 + 18 = \frac{304}{3} \approx 101.33\).

2. The probability of waiting exactly 10 minutes:

Since \(Y\) is continuous, \(P(Y = 10) = 0\).

3. The probability of waiting at most 5 minutes:

\(
P(Y \leq 5) = \int_{0}^{5} \frac{1}{10}y \, dy = \frac{1}{10} \int_{0}^{5} y \, dy = \frac{1}{10} \cdot \frac{25}{2} = \frac{5}{4}.
\)

 4. The IQR:
Find the quartiles \(q_1\) and \(q_3\) by solving:
\(
\int_{0}^{q_1} f_Y(y) \, dy = 0.25, \quad \int_{0}^{q_3} f_Y(y) \, dy = 0.75.
\)
For \(q_1\):
\(
\int_{0}^{q_1} \frac{1}{10}y \, dy = 0.25 \implies \frac{1}{10} \cdot \frac{q_1^2}{2} = 0.25 \implies q_1 = \sqrt{5}.
\)

For \(q_3\), note that it lies in the second interval \(10 \leq q_3 \leq 30\):
\(
\int_{0}^{10} \frac{1}{10}y \, dy + \int_{10}^{q_3} \frac{1}{10} \, dy = 0.75.
\)
\(
\frac{1}{10} \cdot \frac{10^2}{2} + \frac{1}{10}(q_3 - 10) = 0.75 \implies 5 + \frac{q_3 - 10}{10} = 0.75.
\)
Solving for \(q_3\), \(q_3 = 20\).

The IQR is:
\(
q_3 - q_1 = 20 - \sqrt{5}.
\)
 
 \vspace{2cm}
 
 **c. Bus schedule to produce \(k = 5\) and generalization for \(k \geq 1\). **
  
 Ans: For \(k = 5\), divide the hour into 5 equal intervals: \(0, 12, 24, 36, 48, 60\). The PDF will have 5 stair-steps.

To generalize for any \(k \geq 1\), the bus schedule divides the hour into \(k\) equal intervals:
\(
0, \frac{60}{k}, \frac{120}{k}, \dots, 60.
\)
The PDF for \(Y\) will have \(k\) stair-steps, with slopes proportional to the interval length.
 
 \vspace{2cm}
 
 
 


